---
title: "p8105_hw6_CY2772"
author: "Chenhui Yan"
date: "2024-12-02"
output: github_document
---

```{r}
library(dplyr)
library(tidyverse)
library(broom)
library(ggplot2)
library(forcats) 
library(rsample) 
library(modelr)
```

# Problem 2
## 2a) Data cleaning
```{r}

homicide = read_csv("./homicide-data.csv") %>%
  # Replace "Unknown" with NA in victim_age
  mutate(victim_age = na_if(victim_age, "Unknown")) %>%
  # Unite city and state
  unite(col = "city_state", city, state, sep = ", ", remove = TRUE) %>%
  # Continue with data cleaning
  mutate(
    solved_bin = if_else(disposition == "Closed by arrest", 1, 0),
    victim_age = parse_number(victim_age),
    victim_race = factor(victim_race, levels = c("White", "Black")),
    victim_sex = factor(victim_sex)
  ) %>%
  # Apply filters
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
  )

homicide = homicide %>%
  filter(!is.na(victim_age))


```
## 2b1) Define the Logistic Regression Function
```{r}
# Define the Logistic Regression Function
fit_logistic_regression <- function(data) {
  model = glm(
    solved_bin ~ victim_age + victim_race + victim_sex,
    family = binomial(),
    data = data
  )
  
  # Tidy the model output with exponentiated estimates and confidence intervals
  tidy_model = tidy(model, exponentiate = TRUE, conf.int = TRUE)
  
  # Extract the row corresponding to 'victim_sexMale'
  sex_term = tidy_model %>%
    filter(term == "victim_sexMale")
  
  # Check if 'victim_sexMale' exists in the model
  if (nrow(sex_term) == 1) {
    result <- sex_term %>%
      select(
        odds_ratio = estimate,
        conf_lower_95 = conf.low,
        conf_upper_95 = conf.high
      )
  } else {
   
    result <- tibble(
      odds_ratio = NA_real_,
      conf_lower_95 = NA_real_,
      conf_upper_95 = NA_real_
    )
  }
  
  return(result)
}

```

## 2b) Glm for Baltimore, MD
```{r}
baltimore = homicide %>%
  filter(city_state == "Baltimore, MD") %>%
  mutate(
    victim_age = as.numeric(victim_age),
    victim_race = factor(victim_race),
    victim_sex = factor(victim_sex)
  )

# Fit the logistic regression model
baltimore_glm <- glm(
  solved_bin ~ victim_age + victim_race + victim_sex,
  data = baltimore,
  family = binomial()
)

# Get tidy results with exponentiated estimates and confidence intervals
baltimore_results <- tidy(
  baltimore_glm,
  exponentiate = TRUE,
  conf.int = TRUE
) %>%
  filter(term == "victim_sexMale") %>%
  select(
    term,
    odds_ratio = estimate,
    conf_lower_95 = conf.low,
    conf_upper_95 = conf.high
  )

# Display the results
print(baltimore_results)

```
## 2C) Apply for every city
```{r}
# Apply the Logistic Regression Function Across Cities
city_or_results <- homicide %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    glm_results = map(data, fit_logistic_regression)
  ) %>%
  unnest(cols = glm_results) %>%
  select(city_state, odds_ratio, conf_lower_95, conf_upper_95)

# Clean and Prepare the Final Statistics
homicide_stat <- city_or_results %>%
  mutate(
    city_state = as.factor(city_state)
  ) %>%
  rename(
    "Log_OR" = "odds_ratio",
    "CI_Lower_95" = "conf_lower_95",
    "CI_Upper_95" = "conf_upper_95"
  ) %>%
  drop_na(Log_OR, CI_Lower_95, CI_Upper_95)  # Optional: Remove rows with NA values

# Display the Results as a Neatly Formatted Table
knitr::kable(homicide_stat, 
             caption = "Adjusted Odds Ratios for Solving Homicides by Victim Sex Across Cities",
             digits = 3)
```
## 2d) Plot the result
```{r}
# Reorder 'city_state' based on Log_OR for better visualization
homicide_stat <- homicide_stat %>%
  mutate(
    city_state = fct_reorder(city_state, 	Log_OR)
  )

# Create the Plot
ggplot(homicide_stat, aes(x = city_state, y = 	Log_OR, color = city_state)) +
  geom_point(size = 3) +  # Increase point size for better visibility
  geom_errorbar(aes(ymin = CI_Lower_95	, ymax = CI_Upper_95), width = 0.2) +
  coord_flip() +  # Flip coordinates for horizontal alignment
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides by City",
    subtitle = "Comparing Male vs Female Victims",
    x = "City",
    y = "Odds Ratio (Male vs Female Victims)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",                   
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red")  # Reference line at OR = 1

```

The plot illustrates how gender influences homicide resolution across different cities. An odds ratio (OR) above 1 means cases with male victims are more likely to be solved than those with female victims. Cities like Albuquerque, NM, and Stockton, CA, have wide confidence intervals (CIs), indicating uncertainty due to smaller sample sizes. When a CI does not include 1, the effect is statistically significant. Conversely, cities with CIs that cross 1, such as Atlanta, GA, show no significant gender-based difference in resolution likelihood

# Problem 3
## 1a) data cleaning
```{r}
birth_data = read_csv("./birthweight.csv")
str(birth_data)
birth_data = birth_data %>% 
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    frace = factor(frace),
    mrace = factor(mrace),
    malform = factor(malform)
  ) %>% 
  drop_na()
summary(birth_data)
```
## 2) Propose and Fit a Regression Model for Birthweight
```{r}
# Fit the initial regression model
model_initial = lm(bwt ~ gaweeks + ppbmi + smoken + malform + babysex + frace + mrace + momage + wtgain, 
                    data = birth_data)

# Summarize the model
summary(model_initial)

```
```{r}
# Add predictions and residuals to the dataset
birthweight = birth_data %>%
  add_predictions(model_initial, var = "pred") %>%
  add_residuals(model_initial, var = "resid")

# Plot residuals vs. fitted values
ggplot(birthweight, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Residuals vs. Fitted Values",
    x = "Fitted Values (Predicted Birthweight)",
    y = "Residuals"
  )

```

## 3a) Compare the Model to Two Others
```{r}
# Model 1: Using length at birth and gestational age as predictors (main effects only).
model1 <- lm(bwt ~ blength + gaweeks, data = birthweight)
# Model 2: Using head circumference, length, sex, and all interactions (including the three-way interaction).
model2 <- lm(bwt ~ bhead * blength * babysex, data = birthweight)

```
## 3b) Set Up Cross-Validation
```{r}
# Set seed for reproducibility
set.seed(123)

# Generate Monte Carlo cross-validation splits
cv_splits = mc_cv(birthweight, times = 10, prop = 0.8)
# Function to compute RMSE for a given model and test data
compute_rmse = function(model, test_data) {
  predictions = predict(model, newdata = test_data)
  sqrt(mean((test_data$bwt - predictions)^2))
}
# Function to fit models and compute RMSEs for one split
compute_rmse_split = function(split) {
  train_data = analysis(split)
  test_data = assessment(split)
  
  # Fit models on training data
  model_custom_cv = lm(bwt ~ gaweeks + babysex + momage + wtgain + smoken +
                          parity + malform + ppbmi, data = train_data)
  model1_cv = lm(bwt ~ blength + gaweeks, data = train_data)
  model2_cv = lm(bwt ~ bhead * blength * babysex, data = train_data)
  
  # Compute RMSEs on test data
  tibble(
    rmse_custom = compute_rmse(model_custom_cv, test_data),
    rmse_model1 = compute_rmse(model1_cv, test_data),
    rmse_model2 = compute_rmse(model2_cv, test_data)
  )
}
# Apply the function to all splits and collect RMSEs
rmse_results = cv_splits %>%
  mutate(rmse = map(splits, compute_rmse_split)) %>%
  unnest(rmse)
# Calculate mean RMSEs for each model
rmse_summary <- rmse_results %>%
  summarise(
    mean_rmse_custom = mean(rmse_custom),
    mean_rmse_model1 = mean(rmse_model1),
    mean_rmse_model2 = mean(rmse_model2)
  )

print(rmse_summary)

```

```{r}
rmse_long = rmse_results %>%
  pivot_longer(
    cols = starts_with("rmse_"),
    names_to = "model",
    values_to = "rmse"
  )
# Violin plot for RMSE distribution
ggplot(rmse_long, aes(x = model, y = rmse)) +
  geom_violin() +
  labs(
    title = "Model Comparison via RMSE",
    x = "Model",
    y = "RMSE"
  ) +
  theme_minimal()
# Plot RMSE distributions for each model
ggplot(rmse_long, aes(x = model, y = rmse)) +
  geom_boxplot() +
  labs(
    title = "Cross-Validated RMSE for Different Models",
    x = "Model",
    y = "RMSE"
  ) +
  scale_x_discrete(labels = c("Custom Model", "Model 1", "Model 2"))

```

As a result, the model using head circumference, length, sex, and all interactions (including the three-way interaction) between these variables is optimal.
