---
title: "p8105_hw6_CY2772"
author: "Chenhui Yan"
date: "2024-12-02"
output: github_document
---
```{r}
library(dplyr)
library(tidyverse)
library(broom)
library(ggplot2)
library(forcats) 
```

# Problem 2
## 2a) Data cleaning
```{r}

homicide = read_csv("./homicide-data.csv") %>%
  # Replace "Unknown" with NA in victim_age
  mutate(victim_age = na_if(victim_age, "Unknown")) %>%
  # Unite city and state
  unite(col = "city_state", city, state, sep = ", ", remove = TRUE) %>%
  # Continue with data cleaning
  mutate(
    solved_bin = if_else(disposition == "Closed by arrest", 1, 0),
    victim_age = parse_number(victim_age),
    victim_race = factor(victim_race, levels = c("White", "Black")),
    victim_sex = factor(victim_sex)
  ) %>%
  # Apply filters
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
  )

homicide = homicide %>%
  filter(!is.na(victim_age))


```
## 2b1) Define the Logistic Regression Function
```{r}
# Define the Logistic Regression Function
fit_logistic_regression <- function(data) {
  model = glm(
    solved_bin ~ victim_age + victim_race + victim_sex,
    family = binomial(),
    data = data
  )
  
  # Tidy the model output with exponentiated estimates and confidence intervals
  tidy_model = tidy(model, exponentiate = TRUE, conf.int = TRUE)
  
  # Extract the row corresponding to 'victim_sexMale'
  sex_term = tidy_model %>%
    filter(term == "victim_sexMale")
  
  # Check if 'victim_sexMale' exists in the model
  if (nrow(sex_term) == 1) {
    result <- sex_term %>%
      select(
        odds_ratio = estimate,
        conf_lower_95 = conf.low,
        conf_upper_95 = conf.high
      )
  } else {
   
    result <- tibble(
      odds_ratio = NA_real_,
      conf_lower_95 = NA_real_,
      conf_upper_95 = NA_real_
    )
  }
  
  return(result)
}

```

## 2b) Glm for Baltimore, MD
```{r}
baltimore = homicide %>%
  filter(city_state == "Baltimore, MD") %>%
  mutate(
    victim_age = as.numeric(victim_age),
    victim_race = factor(victim_race),
    victim_sex = factor(victim_sex)
  )

# Fit the logistic regression model
baltimore_glm <- glm(
  solved_bin ~ victim_age + victim_race + victim_sex,
  data = baltimore,
  family = binomial()
)

# Get tidy results with exponentiated estimates and confidence intervals
baltimore_results <- tidy(
  baltimore_glm,
  exponentiate = TRUE,
  conf.int = TRUE
) %>%
  filter(term == "victim_sexMale") %>%
  select(
    term,
    odds_ratio = estimate,
    conf_lower_95 = conf.low,
    conf_upper_95 = conf.high
  )

# Display the results
print(baltimore_results)

```
## 2C) Apply for every city
```{r}
# Apply the Logistic Regression Function Across Cities
city_or_results <- homicide %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    glm_results = map(data, fit_logistic_regression)
  ) %>%
  unnest(cols = glm_results) %>%
  select(city_state, odds_ratio, conf_lower_95, conf_upper_95)

# Clean and Prepare the Final Statistics
homicide_stat <- city_or_results %>%
  mutate(
    city_state = as.factor(city_state)
  ) %>%
  rename(
    "Log_OR" = "odds_ratio",
    "CI_Lower_95" = "conf_lower_95",
    "CI_Upper_95" = "conf_upper_95"
  ) %>%
  drop_na(Log_OR, CI_Lower_95, CI_Upper_95)  # Optional: Remove rows with NA values

# Display the Results as a Neatly Formatted Table
knitr::kable(homicide_stat, 
             caption = "Adjusted Odds Ratios for Solving Homicides by Victim Sex Across Cities",
             digits = 3)
```
## 2d) Plot the result
```{r}
# Reorder 'city_state' based on Log_OR for better visualization
homicide_stat <- homicide_stat %>%
  mutate(
    city_state = fct_reorder(city_state, 	Log_OR)
  )

# Create the Plot
ggplot(homicide_stat, aes(x = city_state, y = 	Log_OR, color = city_state)) +
  geom_point(size = 3) +  # Increase point size for better visibility
  geom_errorbar(aes(ymin = CI_Lower_95	, ymax = CI_Upper_95), width = 0.2) +
  coord_flip() +  # Flip coordinates for horizontal alignment
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides by City",
    subtitle = "Comparing Male vs Female Victims",
    x = "City",
    y = "Odds Ratio (Male vs Female Victims)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",                   
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red")  # Reference line at OR = 1

```
The plot illustrates how gender influences homicide resolution across different cities. An odds ratio (OR) above 1 means cases with male victims are more likely to be solved than those with female victims. Cities like Albuquerque, NM, and Stockton, CA, have wide confidence intervals (CIs), indicating uncertainty due to smaller sample sizes. When a CI does not include 1, the effect is statistically significant. Conversely, cities with CIs that cross 1, such as Atlanta, GA, show no significant gender-based difference in resolution likelihood




